% !TEX root = proposal.tex
\chapter{Introduction}

{\bf Thesis Statement} Modern applications can benefit from read and
write interaction with a network's control-plane, yet such interaction is
currently unsupported due to security and fairness concerns. We design
a practical and feasible platform that addresses these challenges,
and demonstrate its benefits. In addition, we can support a network's
administrator in the safe and effective use of such a platform.

\vskip 1em

\section{The Need for Delegated Network Management}

Today's applications, whether running in datacenters, enterprise, campus, or
home networks have an increasingly difficult relationship with the
network. Networks are the shared fabric interconnecting users, applications,
and devices, and fluctuating, unpredictable network performance and reliability
create challenges and uncertainty for network administers, application
developers, and frustrated end-users alike. As a result, software developers,
researchers, and administrators expend considerable effort to work \emph{around}
the network rather than work \emph{with} the network: video conferencing
applications constantly probe network bandwidth~\cite{SkypeTraffic}, overlay networks are used to
re-route traffic~\cite{SkypeTraffic}, network paths are reactively reconfigured based on inferences~\cite{alfares10hedera},
and humans are required to throttle heavy network loads in response to planned
or unplanned shifts in traffic matrices.
%
Using humans for network control, however, is no panacea, having
been responsible for significant recent outages at both
Github~\cite{githubOutage} and
Amazon~\cite{awsOutage}.

At a minimum, packet networks forward data, collect traffic statistics, and
divide traffic based on addresses or other header fields. Increasingly, modern
networks also provide additional services, often implemented via middleboxes,
such as firewalling, compression, encryption, threat-detection, acceleration, and caching.
Yet, all of these features are, for the most part, invisible to the applications
passing traffic through them, or only available via rudimentary interfaces
such as DSCP header bits.

With greater visibility into and control of the network's state,
a conferencing application could request bandwidth for a video call,
and learn via the network that while only a guaranteed audio call is available
now, it could reserve a video call in one hour.
An intrusion detection script on a user machine could request that the
\emph{network} filter traffic from a specific source.
An important RPC service could protect latency-sensitive flows from competing
background traffic.
%An RPC service could inform the network that its flows will be short, but with a
%deadline, which could enable shortest-deadline-first scheduling at
%the switches~\cite{wilson11d3}.
Or, a MapReduce-style application could request
bandwidth guarantees or maximally disjoint paths to improve performance of its shuffle phase.

Such examples suggest that an API should exist between the network's
control-plane and its users, applications, and end-hosts. These
principals need both \emph{read} access to learn the network's
present and potential future conditions, and \emph{write} access to
make independent configuration changes for their own benefit, and
provide helpful knowledge such as future traffic demands, directly
to the network's control-plane.

In this thesis, we introduce the concept of \emph{participatory
networking}, in which the network provides
a configuration API to its users, applications, and end-hosts, and
present the design, implementation, and evaluation of the first
practical participatory networking controller for an
OpenFlow-enabled software-defined network.

In the absence of security guarantees and limited authorities,
participatory networks would be places of anarchy. To be usable,
such networks must provide isolation for their independent
principals, preventing malicious users from hogging bandwidth,
dropping packets, or worse. While isolation could be provided
through network virtualization, we believe that is not always the
right abstraction as it hides the fundamentally shared aspect of
networks.  

% new
By contrast, participatory networks reveal the conflicts between
their principals and directly expose the control-plane's view of
the network. Principals can use this greater awareness to make
better-informed decisions about how to use or reconfigure the
network.

% new

\section{Software-Defined Networking}

The recent development of Software-Defined Networks (SDNs)
%which provide a logically centralized abstraction of network
%control,
%built on programmable devices,
offers a platform to realize the vision of participatory networks~\cite{Greenberg:2005,McKeown:2008}.
SDNs separate
the logic that controls the network from its physical
devices, allowing configuration \emph{programs} to operate on a
high-level, global, and consistent view of the network. 
%  allowing network-wide policies to be determined
% centrally~\cite{Casado:2009, Gude:2008}, and expressed at a higher
% level of abstraction than configuration of individual network
% elements~\cite{Foster:2010,Hinrichs:2009, Kim:2010}
This change has brought significant advances to datacenter and
enterprise networks, such as high-level
specification of access control~\cite{Nayak:2009} and
QoS~\cite{Kim:2010}, safe experimentation~\cite{Sherwood:2010},
in-network load-balancing~\cite{wang11wild}, and seamless VM 
migration~\cite{Erickson:2008}.

In a traditional network, the control-plane, which determines how
to transmit packets is distributed across the network's switches. Each
switch independently runs distributed algorithms (\eg, Bellman-Ford) to determine the
packet forwarding policy. By contrast, an SDN has a logically centralized
controller, running on a commodity server platform, which calculates
a forwarding policy (\eg, with Dijkstra's algorithm). The controller implements
this policy by programming the switches' data-planes, which process
traffic at line-rate.

This programming is accomplished via a protocol such as
OpenFlow~\cite{McKeown:2008}, which provides a common
abstraction of a switch's data-plane. The principal component of this
abstraction is a sequence of prioritized ``Match-Action'' pairs, where
matches list a set of packet header fields, and actions specify handling
of the matching packets: transmit via a port, place in a queue, drop, etc.

\section{A Prototype System}

We have developed a prototype OpenFlow-based controller
for participatory networks called \sys.
The \sys controller implements a capability system --
it delegates read and
write authority, with optional restrictions, from the network's
administrators to the users, or applications and hosts acting on
their behalf.  The controller is logically centralized, has a global
view of the network, and implements the principals' high-level
intents by changing the configuration of the actual network devices.
In addition, we implement and evaluate the examples described
above by augmenting four real applications (\xref{sec:Evaluation}).

\sys's user-facing API serves as the next layer on the current SDN stack.
The abstractions provided by software-defined networks allow us
to reason formally about \sys's design, and ensure the network
continues to provide baseline levels of fairness and security, even as principals
dynamically invoke their capabilities.

Our design addresses the two key challenges of participatory networks:
how to safely decompose control and visibility of the network, and how to resolve conflicts
between participants and across requests. 
\sys's solutions to these challenges were developed both through formal reasoning,
and by porting real-world applications to solve existing use cases.
We start with an overview of our solution in \xref{sec:overview},
followed by more in-depth discussions for each challenge in \xref{sec:delegation}
and \xref{sec:conflicts} respectively.


\vskip 1em

Many approaches to achieving some of these goals have been previously
proposed including active networking~\cite{ActiveNetworking}, IntServ networking~\cite{rfc1633}, and
distributed reservation protocols such as RSVP~\cite{rfc2205} and NSIS~\cite{rfc5974}.
We discuss their relation to participatory networking in \xref{sec:related-work}.
\sys does not introduce any new
functionalities into the network. Instead, it exposes existing functionalities and
provides a single platform for reasoning about their use. We argue that
this approach provides several advantages: a single target for application
developers, a unified management framework for network administrators,
and, most importantly, the ability to reason across all network resources.

\sys is designed for networks within a single administrative domain
including corporate WANs, datacenters, campus or enterprise
networks, and home networks. In such networks, there is, logically,
a single owner from which authority can be delegated.  \sys's design
does not rely on changes to the end-hosts, such as the use of a
particular hypervisor, making it suitable for networks with
user-owned or managed devices.  

\begin{comment}

\vskip 0.5em

In summary, this thesis makes the following contributions:

\begin{enumerate}

%\item {\color{red}We develop a uniform abstraction to enable this ``conversation,'' and design
%a corresponding API.}

\item We implement a fully-functioning SDN controller which allows a
network's administrators to safely delegate their authority using our API.

\item We analyze a previously proposed algorithm for consolidating
hierarchical policies, and propose a new algorithm that reduces the complexity
from exponential to polynomial.

\item We demonstrate our system's usefulness and
practicality on a real OpenFlow testbed using 
microbenchmarks and four real applications enhanced with our API.


\end{enumerate}

\end{comment}

\section{Proposed Work}

{\color{red}
Briefly discuss the proposed work beyond this...
}